Hydras, Ordinals \& Co. \\ A library in Coq of entertaining formal mathematics

### 2021-10-05 doc/JFLA2022/paper.tex

\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

@@

 Prior stages of this project have already been presented at
 JFLA~\cite{PCiota, JFLA2018paper}.
We present recent evolutions of the library: new results, interaction within the \community project~\cite{CoqCommunity}, and documentation generated with \alectr~\cite{alectryonpaper, alectryongithub}.

@@

\section{Integration within \community}

@@

\community is an informal organization run by volunteer \coq users with objectives to maintain interesting \coq projects in the long-term and help \coq users to collaborate on documentation, tooling, etc.

@@

In the case of \coq, this problem may be even more prevalent as many packages are created by PhD students, or by researchers for a specific paper, but their authors do not intend to maintain the package in the long-term.
%
The authors are however generally very open to having someone else who expresses interest in their work continue the maintenance.
%
\community makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).

Today, \community hosts over 50 projects maintained by over 30 maintainers.
%
The hosted projects come from a variety of origins.
%
Some had been maintained in the past by the \coq development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new \coq versions were conducted.
%
Some were still maintained by their authors but they wanted help to keep them up-to-date.
%
Others were simply left unmaintained and were revived by their adoption in \community.

\subsection{Interactions with other \community projects}

Given the objectives of \community, it makes sense to propose a transfer anytime we encounter an interesting project that is insufficiently maintained.
%
After the project is adopted, large changes and refactorings are explicitly allowed.
%
This means, for instance, that we can consolidate libraries together or split them into several packages if it makes sense.

In order to prove formally that the length of the considered
kind of battles is not given by a primitive recursive function, we used the formalization of primitive recursive functions, part
of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem~\cite{OConnor05, Goedel}.
For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \community.

@@

We plan to extend our libraries in two main directions.
The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710}. This library is dedicated  to the implementation in \coq (SSreflect) of books from  Bourbaki's Elements of Mathematics. It contains data structures which are compatible with our ordinal notations. 
Our plan is to build a  bridge between  the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.
Another direction would be to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.

>>>>

\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}
\newcommand{\Hydras}{Hydras \& Co.\xspace}
\newcommand{\Coqcommunity}{Coq-community\xspace}

@@

 Prior stages of this project have already been presented at
 JFLA~\cite{PCiota, JFLA2018paper}.

We present recent evolutions of the library: new results, interaction within the \community project~\cite{CoqCommunity}, and documentation generated with \alectr~\cite{alectryonpaper, alectryongithub}.

@@

\section{Integration within \Coqcommunity}

@@

\Coqcommunity is an informal organization run by volunteer Coq users with objectives to maintain interesting Coq projects in the long-term and help Coq users to collaborate on documentation, tooling, etc.

@@

In the case of \coq, this problem may be even more prevalent as many packages are created by PhD students, or by researchers for a specific paper, but their authors do not intend to maintain the package in the long-term.
%
The authors are however generally very open to having someone else who expresses interest in their work continue the maintenance.
%

\Coqcommunity makes that easy by providing a process for transferring or forking an unmaintained package, tooling for setting up good maintenance practices (such as continuous integration) and by making it possible for someone to take over a package without committing in the long-term (as maintainers who drop out can easily be replaced by some other volunteers).

Today, \Coqcommunity hosts over 50 projects maintained by over 30 maintainers.
%
The hosted projects come from a variety of origins.
%
Some had been maintained in the past by the \coq development team on behalf of the authors, but this meant that only minimal changes required to make the project build with new \coq versions were conducted.
%
Some were still maintained by their authors but they wanted help to keep them up-to-date.
%
Others were simply left unmaintained and were revived by their adoption in \Coqcommunity.

\subsection{Interactions with other \Coqcommunity projects}

Given the objectives of \Coqcommunity, it makes sense to propose a transfer anytime we encounter an interesting project that is insufficiently maintained.
%
After the project is adopted, large changes and refactorings are explicitly allowed.
%
This means, for instance, that we can consolidate libraries together or split them into several packages if it makes sense.

In order to prove formally that the length of the considered
kind of battles is not given by a primitive recursive function, we used the formalization of primitive recursive functions, part
of Russell O'Connor's contribution on G\"{o}del first incompleteness theorem~\cite{OConnor05, Goedel}.
For this purpose, and above all by consideration of the scientific interest of this contribution, we chose to host and maintain this work within \Coqcommunity.

@@

We plan to extend our libraries in two main directions.

The \gaia project~\cite{Gaia}, also maintained on \community, contains a development by José Grimm~\cite{grimm:hal-00911710}. This library is dedicated  to the implementation in \coq (SSreflect) of books from  Bourbaki's Elements of Mathematics. It contains data structures which are compatible with our ordinal notations. 
Our plan is to build a  bridge between  the combinatorial results of \Hydras and the set-theoretic content of \gaia, and make possible the transfer of theorems between both libraries.

Another direction would be to write in \coq a formal proof of the original statement of Theorem~\ref{kp:thm2}, using O'Connor's formalization of Peano Arithmetic~\cite{Goedel}.
